---
title: "Languages in Canada in 2016"
author: "Henry Truong"
date: "17/11/2021"
output: 
  ioslides_presentation:
    keep_md: yes
---

```{r library, include = F}
library(tidyverse)
library(ggplot2)
library(plotly)
```


## Introduction

```{r, include = F}
can_lang = read_csv(file = 'https://raw.githubusercontent.com/UBC-DSCI/introduction-to-datascience/master/data/can_lang.csv')
```

The data set is related to languages spoken by Canadian residents and collected 
during the 2016 Canadian census. The data can be downloaded from [this link](https://raw.githubusercontent.com/UBC-DSCI/introduction-to-datascience/master/data/can_lang.csv).  

```{r, eval = F}
can_lang = read_csv(file = fileUrl)
dim(can_lang)
```

```{r, echo = F, comment = ""}
dim(can_lang)
```

There are 214 languages, each of which has 6 properties: (1) `category`, (2) `language`, (3) `mother_tongue`, (4) `most_at_home`, (5) `most_at_work`, (6) `lang_known`  

The application is used to answer 2 questions


## Question 1

#### What are the top 3-10 languages of the selected category spoken most as mother tongue?

```{r, echo = F, fig.align = 'center', fig.cap = 'Top 3 Aboriginal Languages spoken as Mother Tongue'}
abo_lang = filter(can_lang, category == 'Aboriginal languages')
arranged_abo_lang = arrange(abo_lang, by = desc(mother_tongue))
top5_abo_lang = slice(arranged_abo_lang, 1:5)
ggplot(top5_abo_lang, aes(x = mother_tongue, 
                          y = reorder(language, mother_tongue))) +
  geom_bar(stat = "identity") +
  xlab("Mother Tongue (Number of Canadian Residents)") +
  ylab("Language") + 
  theme_bw() + 
  theme(axis.text.x = element_text(color = "grey40", size = 15, hjust = .5, vjust = .5),
        axis.text.y = element_text(color = "grey40", size = 15, hjust = 1, vjust = .5),
        axis.title.x = element_text(color = "grey20", size = 18, hjust = .5, vjust = 0),
        axis.title.y = element_text(color = "grey20", size = 18, hjust = .5, vjust = 1))
```


## Question 2

#### What is the number of people speaking a language at work, given the number of them speaking it at home?  

```{r, eval = F}
# Extract only top 20 Aboriginal languages spoken as mother tongue
top20_abo_lang = slice(arranged_abo_lang, 1:20)
# Build model
mdl_abo = lm(most_at_work ~ most_at_home, data = top20_abo_lang)
# Predict most_at_work value from most_at_home value of 20000
homeNew = 20000
workNew_abo = predict(mdl_abo, newdata = data.frame(most_at_home = homeNew))
# Create scatter plot and add linear line and predicted point 
with(top20_abo_lang, plot(x = most_at_home, y = most_at_work,
                          xlab = 'Home', ylab = 'Work',
                          xlim = c(0, 40000), ylim = c(0, 10000),
                          bty = 'n', pch = 16))
abline(mdl_abo, lwd = 3, col = 'red')
points(homeNew, workNew_abo, cex = 1.5, pch = 16, col = 'red')
```

## Question 2

```{r, echo = F, fig.align = 'center'}
# Extract only top 20 Aboriginal languages spoken as mother tongue
top20_abo_lang = slice(arranged_abo_lang, 1:20)
# Build model
mdl_abo = lm(most_at_work ~ most_at_home, data = top20_abo_lang)
# Predict most_at_work value from most_at_home value of 20000
homeNew = 20000
workNew_abo = predict(mdl_abo, newdata = data.frame(most_at_home = homeNew))
# Create scatter plot and add linear line and predicted point 
with(top20_abo_lang, plot(x = most_at_home, y = most_at_work,
                          xlab = 'Home', ylab = 'Work',
                          xlim = c(0, 40000), ylim = c(0, 10000),
                          bty = 'n', pch = 16))
abline(mdl_abo, lwd = 3, col = 'red')
points(homeNew, workNew_abo, cex = 1.5, pch = 16, col = 'red')
abline(h = workNew_abo, lwd = 1, col = 'gray')
abline(v = homeNew, lwd = 1, col = 'gray')
```

The predicted value of `most_at_work` is `r round(workNew_abo)` when `most_at_home` is 20000


